---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rjd3sts


## Installation

Running rjd3 packages requires Java 17 or higher. How to set up such a configuration in R is explained [here](https://jdemetra-new-documentation.netlify.app/#installing-the-software).

To get the current stable version (from the latest release):

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rjdemetra/rjd3toolkit@*release")
remotes::install_github("rjdemetra/rjd3sts@*release")
```

To get the current development version from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rjdemetra/rjd3sts")
```


## Package Maintenance and contributing

Any contribution is welcome and should be done through pull requests and/or issues.
pull requests should include **updated tests** and **updated documentation**. If functionality is changed, docstrings should be added or updated.


## Licensing

The code of this project is licensed under the [European Union Public Licence (EUPL)](https://joinup.ec.europa.eu/collection/eupl/eupl-text-eupl-12).
