---
title: "Sructural Time Series using JDemetra+"
author: "Jean Palate"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sructural Time Series using JDemetra+}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

```{r echo=FALSE}
suppressPackageStartupMessages(library(rjd3toolkit))
suppressPackageStartupMessages(library(rjd3sts))
library(knitr)
```

# Data

We illustrate the various methods using two sets of time series:

* The **retail** data set contains monthly figures over US retail activity of various categories of goods and services from 1992 to 2010.
* The **ABS** data set contains long monthly series over Australian retail trade.


# Overview of the supported state space forms

# Description of the state blocks

## ar

### Introduction

The auto-regressive block is defined by

$$\Phi\left(B\right)y_t=\epsilon_t $$ 

where:  

$$\Phi\left(B\right)=1+\varphi_1 B + \cdots + \varphi_p B^p $$   

is an auto-regressive polynomial. 

Let $\gamma_i$ be the autocovariances of the process

Using those notations, the state-space block can be written as follows :

### State block: 

$$ \alpha_t= \begin{pmatrix} y_t \\ y_{t-1} \\ \vdots \\ y_{t-p+1} \end{pmatrix}$$  
The state block can be extended with additional lags. That can be useful in complex (multi-variate) models

### Dynamics

$$ T_t = \begin{pmatrix}-\varphi_1 & \cdots & \cdots  & -\varphi_p  \\
                        1          & \cdots &  \cdots &  0          \\ 
						\vdots     & \ddots &  \ddots & \vdots\\ 
						0          &   0    &  1      & 0  \end{pmatrix}$$

$$ S_t = \sigma_{ar} \begin{pmatrix} 1 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  

$$ V_t = S S' $$

### Default loading

$$ Z_t = \begin{pmatrix} 1 & 0 & \cdots & 0\end{pmatrix}$$

### Initialization 

$$ \alpha_{-1} = \begin{pmatrix}0 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  

$$ P_{*} = \Omega $$
$\Omega$ is the unconditional covariance of the state array; it can be easily derived using the MA representation. We have:

$$ \Omega\left(i,0\right) = \gamma_i $$  

$$ \Omega\left(i,j\right) = \Omega\left(i-1,j-1\right)-\psi_i \psi_j $$  

```{r }
b_ar<-ar("ar", c(-.2, .4, -.1), nlags=5)
knit_print(block_t(b_ar))

```
## ar2

### Introduction

An alternative representation of the auto-regressive block will be very useful for the purposes of reflecting expectations. The  process is defined as above:

$$\Phi\left(B\right)y_t=\epsilon_t $$ 

where:  

$$\Phi\left(B\right)=1+\varphi_1 B + \cdots + \varphi_p B^p $$   

is an auto-regressive polynomial. However, modeling data that refers to expectations may require including conditional expectations in 
the state vector. Thus,  the same type of representation that is used for the ARMA model will be considered here.

Let $\gamma_i$ be the autocovariances of the model. We also define the size of our state vector as $r0=max(p,h+1)$, where $h$ is the forecast horizon
desired by the user. If the user needs to use $nlags$ lagged values, whose default value is zero. Then the size of the state vector will be $r=r0+nlags$ 

Using those notations, the state-space model can be written as follows :

### State block: 

$$ \alpha_t= \begin{pmatrix} 
y_{t-nlags} \\ 
\vdots  \\ 
y_{t-1}  \\ 
\hline
y_{t}  \\ 
y_{t+1|t} \\ 
\vdots \\ 
y_{t+h|t} 
\end{pmatrix}$$  

where $y_{t+i|t}$ is the orthogonal projection of
$y_{t+i}$ on the subspace generated by ${y\left(s\right):s \leq t}$. Thus, it is the forecast function with respect to the semi-infinite sample.
We also have that $y_{t+i|t} = \sum_{j=i}^\infty {\psi_j \epsilon_{t+i-j}}$

### Dynamics


$$ T_t = \begin{pmatrix} 0 &1 & 0 & \cdots & 0  \\0& 0 & 1 & \cdots & 0\\ \vdots & \vdots & \vdots & \ddots & \vdots\\ 0 & 0 & 0 & \cdots & 1\\
-\varphi_r & \cdots  & \cdots & \cdots &-\varphi_1 \end{pmatrix}$$

<p style="text-align: center;"> with $\varphi_{j}=0$ for $j>p$  </p>

$$ S_t = \sigma_{ar} \begin{pmatrix} 0 \\  
\vdots \\
0\\
\hline
1 \\ \psi_1 \\ \vdots\\ \psi_s \end{pmatrix} $$  

$$ V_t = S S' $$



### Default loading

$$ Z_t = \begin{pmatrix} 0 & \cdots  &0  & | & 1 & 0 & \cdots & 0\end{pmatrix}$$


### Initialization 

$$ \alpha_{-1}  = \begin{pmatrix} 0 \\  
\vdots \\
0\\
\hline
0 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  



$$ P_{*} = \Omega $$

$\Omega$ is the unconditional covariance of the state array; it can be easily derived using the MA representation. We have:

$$ \Omega\left(i,0\right) = \gamma_i $$  

$$ \Omega\left(i,j\right) = \Omega\left(i-1,j-1\right)-\psi_i \psi_j $$  

```{r }
b_ar2<-ar2("ar2", c(-.2, .4, -.1), nlags=3, nfcasts=2)
knit_print(block_t(b_ar2))

```

## arma

### Introduction

The arma block is defined by

$$\Phi\left(B\right)y_t=\Theta\left(B\right)\epsilon_t $$ 

where:  

$$\Phi\left(B\right)=1+\varphi_1 B + \cdots + \varphi_p B^p $$   

$$\Theta\left(B\right)=1+\theta_1 B + \cdots + \theta_q B^q $$    

are the auto-regressive and the moving average polynomials. 

The MA representation of the process is $y_t=\sum_{i=0}^\infty {\psi_i \epsilon_{t-i}}$. Let $\gamma_i$ be the autocovariances of the model. We also define: $r=\max\left(p, q+1\right), \quad s=r-1$.

Using those notations, the state-space block can be written as follows :

### State block: 

$$ \alpha_t= \begin{pmatrix} y_t \\ y_{t+1|t} \\ \vdots \\ y_{t+s|t} \end{pmatrix}$$  

where $y_{t+i|t}$ is the orthogonal projection of
$y_{t+i}$ on the subspace generated by ${y\left(s\right):s \leq t}$.Thus, it is the forecast function with respect to the semi-infinite sample.
We also have that $y_{t+i|t} = \sum_{j=i}^\infty {\psi_j \epsilon_{t+i-j}}$

### Dynamics

$$ T_t = \begin{pmatrix}0 &1 & 0 & \cdots & 0  \\0& 0 & 1 & \cdots & 0\\ \vdots & \vdots & \vdots & \ddots & \vdots\\ 0 & 0 & 0 & \cdots & 1\\
-\varphi_r & \cdots  & \cdots & \cdots &-\varphi_1 \end{pmatrix}$$

<p style="text-align: center;"> with $\varphi_{j}=0$ for $j>p$  </p>


$$ S_t = \begin{pmatrix}1 \\ \psi_1 \\ \vdots\\ \psi_s \end{pmatrix} $$  

$$ V_t = S S' $$

### Default loading

$$ Z_t = \begin{pmatrix} 1 & 0 & \cdots & 0\end{pmatrix}$$

### Initialization 

$$ \alpha_{-1} = \begin{pmatrix}0 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  

$$ P_{*} = \Omega $$

$\Omega$ is the unconditional covariance of the state array; it can be easily derived using the MA representation. We have:

$$ \Omega\left(i,0\right) = \gamma_i $$  

$$ \Omega\left(i,j\right) = \Omega\left(i-1,j-1\right)-\psi_i \psi_j $$  

```{r }
b_arma<-arma("arma", ar=c(-.2, .4, -.1), ma=c(.3, .6))
knit_print(block_t(b_arma))
knit_print(block_p0(b_arma))

```

## arima



# Examples

## [Basic structural models](bsm.html)

## [Time varying trading days](tdvar.html)

## [Regular period cubic splines](regsplines.html)

## [Time series with a sampling error](regcmp_6.html)


