---
title: "Sructural Time Series using JDemetra+"
author: "Jean Palate"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sructural Time Series using JDemetra+}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

```{r echo=FALSE}
suppressPackageStartupMessages(library(rjd3toolkit))
suppressPackageStartupMessages(library(rjd3sts))
library(knitr)
```

# Data

We illustrate the various methods using two sets of time series:

* The **retail** data set contains monthly figures over US retail activity of various categories of goods and services from 1992 to 2010.
* The **ABS** data set contains long monthly series over Australian retail trade.


# Overview of the supported state space forms

# Description of the state blocks

## ar

### Introduction

The AR process is defined by

$$\Phi\left(B\right)y_t=\epsilon_t $$ 

where:  

$$\Phi\left(B\right)=1+\varphi_1 B + \cdots + \varphi_p B^p $$   

is an auto-regressive polynomial. 

Let $\gamma_i$ be the autocovariances of the model.

Using those notations, the state-space model can be written as follows :

### State block: 

$$ \alpha_t= \begin{pmatrix} y_t \\ y_{t-1} \\ \vdots \\ y_{t-p+1} \end{pmatrix}$$  
The state block can be extended with additional lags. That can be useful in complex (multi-variate) models

### Dynamics

$$ T_t = \begin{pmatrix}-\varphi_1 & \cdots & \cdots  & -\varphi_p  \\
                        1          & \cdots &  \cdots &  0          \\ 
						\vdots     & \ddots &  \ddots & \vdots\\ 
						0          &   0    &  1      & 0  \end{pmatrix}$$

$$ S_t = \sigma_{ar} \begin{pmatrix} 1 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  

$$ V_t = S S' $$

### Measurement

$$ Z_t = \begin{pmatrix} 1 & 0 & \cdots & 0\end{pmatrix}$$

$$ h_t = 0 $$

### Initialization 

$$ \alpha_{-1} = \begin{pmatrix}0 \\ 0 \\ \vdots\\ 0 \end{pmatrix} $$  

$$ P_{*} = \Omega $$
$\Omega$ is the unconditional covariance of the state array; it can be easily derived using the MA representation. We have:

$$ \Omega\left(i,0\right) = \gamma_i $$  

$$ \Omega\left(i,j\right) = \Omega\left(i-1,j-1\right)-\psi_i \psi_j $$  

```{r }
b_ar<-ar("ar", c(-.2, .4, -.05), nlags=5)
knit_print(block_t(b_ar))

```

# Examples

## [Basic structural models](bsm.html)

## [Time varying trading days](tdvar.html)

## [Regular period cubic splines](regsplines.html)



