---
title: "Time dependent airline model"
author: "Jean Palate"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time dependent airline model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

This document presents some results on time-dependent arima models developed by G. Melard et al., as they are implemented in JDemetra+
The current implementation focuses on airline models.

#### Estimation of time-dependent airline models

The likelihood ratios between time-dependent and normal airline models computed on 62 series of the US-retail trade statistics are presented below. Likelihood ratios above 4 indicate a significant preference for the time-dependent models.

```{r echo=FALSE}
library(knitr)
```


```{r estimation, fig.width=6, fig.height=4}
test<-function(z){
  q<-rjd3sts::tdairline_estimation(z)
  return (q$ltd_sarima$likelihood- q$sarima$likelihood)
}

all<-sapply(rjd3toolkit::Retail, function(z) test(z))

hist(all, breaks=10, main="Distribution of the likelihood ratios")
print(all)

```

#### Canonical decomposition and estimation of the components by means of the Kalman smoother

We present below the seasonal adjustment of a series (retail trade of gasoline stations) and the differences between the series based on the airline model and the time-dependent airline model, using the Kalman smoother
(time-dependent series in red)


```{r decomp}


s<-log(rjd3toolkit::Retail$GasolineStations)

q<-rjd3sts::tdairline_estimation(s)

tdss<-rjd3sts::tdairline_decomposition(s, q$ltd_sarima$parameters)

```

```{r sa, echo=FALSE, fig.width=6, fig.height=4}
th<-q$sarima$parameters[2]
if (th< -.98){th<- -.98}
bth<-q$sarima$parameters[3]
if (bth< -.98){bth<- -.98}

air<-rjd3toolkit::sarima_model(period=12, d=1, theta=th,bd=1, btheta=bth )
ucm<-rjd3toolkit::sarima_decompose(air)

ss<-rjd3toolkit::ucarima_estimate(s, ucm)

sa_td<-ts(s-tdss[,2], frequency = frequency(s), start=start(s))
sa<-ts(s-ss[,2], frequency = frequency(s), start=start(s))

ts.plot(ts.union(sa_td, sa), type='l', col=c('red', 'gray'), main = "SA")

ts.plot(window(ts.union(sa_td, sa), end=1996), type='l', col=c('red', 'gray'), main = "SA (start)")
ts.plot(window(ts.union(sa_td, sa), start=2005), type='l', col=c('red', 'gray'), main = "SA (end)")


s_td<-ts(tdss[,2], frequency = frequency(s), start=start(s))
s<-ts(ss[,2], frequency = frequency(s), start=start(s))

ts.plot(ts.union(s_td, s), type='l', col=c('red', 'gray'), main="Seas")


plot(s_td-s, type='l', main ="Differences between Seas")

t_td<-ts(tdss[,1], frequency = frequency(s), start=start(s))
t<-ts(ss[,1], frequency = frequency(s), start=start(s))

ts.plot(ts.union(t_td, t), type='l', col=c('red', 'gray'), main ="Trend")


```

#### Main results


##### Airline:

log-likelihood = `r q$sarima$likelihood`

$\theta=$ `r q$sarima$parameters[2]`

$\Theta=$ `r q$sarima$parameters[3]`


###### Time-dependent airline

log-likelihood = `r q$ltd_sarima$likelihood`

$\bar{\theta}=$ **`r 0.5*(q$ltd_sarima$parameters[1]+q$ltd_sarima$parameters[2])`** [ `r q$ltd_sarima$parameters[1]`, `r q$ltd_sarima$parameters[2]`]

$\bar{\Theta}=$ **`r 0.5*(q$ltd_sarima$parameters[3]+q$ltd_sarima$parameters[4])`** [ `r q$ltd_sarima$parameters[3]`, `r q$ltd_sarima$parameters[4]`]

###### Canonical decomposition

```{r cd, fig.width=6, fig.height=4}
library("rjd3tramoseats")

airline_decomposition<-function(period, th, bth){
  sarima<-rjd3toolkit::sarima_model("m", period, NULL, 1, th, NULL, 1, bth)
  return(seats_decompose(sarima))
}

airline_variances<-function(period, th, bth){
  ucm<-airline_decomposition(period, th, bth)
  if (is.null(ucm)) return (c(NA, NA, NA)) else return (c(ucm$components[[1]]$var, 
                                                          ucm$components[[2]]$var, 
                                                          ucm$components[[3]]$var))
}

# Gets the variances of the canonical decomposition for airline models with different parameters
th<-q$ltd_sarima$th
bth<-q$ltd_sarima$bth

vars<-sapply(seq(1, length(th)), function(z){return (airline_variances(12,th[z], bth[z]))})
vars<-t(vars)

matplot(vars, main="Models of the components. Innovation variances", type = 'l', ylab="Variances",col=c("red", "blue", "green"))
legend("topleft", legend=c("T", "S", "I"),
       col=c("red", "blue", "green"), lty=1:2, cex=0.8)

```
