---
title: "Time dependent airline model"
author: "Jean Palate"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time dependent airline model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

This document presents some results on time dependent arima models developed by G. Melard, as they are implemented in JDemetra+
The current implementation focuses on airline models.

#### Estimation of time-dependent airline models

The likelihood ratios between time-dependent and normal airline models computed on 63 series of the US-retail trade statistics are presented below. Likelihood ratios above 4 indicate a significant preference for the time-dependent models.

```{r echo=FALSE}
library(knitr)
```


```{r estimation}
test<-function(z){
  q<-rjd3sts::tdairline_estimation(z)
  return (q$ltd_sarima$likelihood- q$sarima$likelihood)
}

all<-sapply(rjd3toolkit::Retail, function(z) test(z))

hist(all, breaks=10)
print(all)

```

#### Canonical decomposition and estimation of the components by means of the Kalman smoother

We present below the seasonal adjustment of a series (total retail trade) and the differences between the sa series based on the airline model and the time-dependent airline model, using the Kalman smoother


```{r sa, fig.width=6, fig.height=4}
s<-log(rjd3toolkit::Retail$RetailAndFoodServicesSalesTotal)

q<-rjd3sts::tdairline_estimation(s)

print('airline')
print(q$sarima$likelihood)
print('tdairline')
print(q$ltd_sarima$likelihood)

tdss<-rjd3sts::tdairline_decomposition(s, q$ltd_sarima$th, q$ltd_sarima$bth)

air<-rjd3toolkit::sarima_model(period=12, d=1, theta=q$sarima$parameters[2],bd=1, btheta=q$sarima$parameters[3] )
ucm<-rjd3toolkit::sarima_decompose(air)

ss<-rjd3toolkit::ucarima_estimate(s, ucm)
plot(s-tdss[,2], type='l', col='red')
lines(s-ss[,2], col='gray')

plot(tdss[,2]-ss[,2], type='l')


```
